### 2022-08-04 latex/PHYS300/PHYS210.tex

\documentclass{article}

\usepackage{../needed}

\title{PHYS210 - HW2}
\author{Alfaifi, Ammar}
\date{\today}

\includeonly{chapters/chapter4}

\begin{document}
    \maketitle

    \paragraph{}
    This is a solution for the problem 2-22.
    We want to find the equation of motion of a charged particle in an  
    electromagnetic field using \textbf{Lorentz equation}.
    If the electric field vector $\vec{E}$ and the magnetic field vector $\vec{B}$,
    the force on a particle of mass $m$ that carries a charge $q$ and has a velocity
    $\vec{v}$ is given
    \begin{equation} 
        \label{eq:lorentz}   
        \vec{F} = q \vec{E} + q \vec{v} \times \vec{B}
    \end{equation}

    where we assume $v \ll c$.
    \paragraph{(a)} If there is no electeic field and the particle enters the
    the magnetic field in direction perpendicular to the magnetic field. The 
    particle follows a uniform circular motion with acceleration of $a_c=m \frac{v^2}{r}$
    where $r$ is the radius of the path of the circular motion.
    Thus, by Newton's second law we have
    $$
        F = ma_c \Rightarrow qvB sin(90^\circ) = m  \frac{v^2}{r}
    $$
    solving for $r$
    $$ r = \frac{mv}{qB} = \frac{v}{\omega_c} $$
    where $\omega_c \equiv qB/m$, is the \textit{cyclotron frequency}.

    \paragraph{(b)} Choose the z-axis to lie in the direction of $\vec{B}$ and let
    the plane containing $\vec{E}$ and $\vec{B}$ be the yz-plane. Thus,
    $$
        \vec{B} = B \hat{z}, \qquad \vec{E} = E_y \hat{y} + E_z \hat{z}
    $$
    Show that the z component of the motion is given by 
    $$
        z(t) = z_0 + \dot{z_0}t + \frac{qE_z}{2m} t^2
    $$
    where $z(0) \equiv z_0$, and $ \dot{z}(0) \equiv \dot{z_0}$.

    Deviding both sides Eq.1, by $q$, we get
    \begin{equation}
        \frac{\vec{F}}{q} = (E_y \hat{y} + E_z \hat{z}) + (v_x \hat{x}+v_y \hat{y}+v_z \hat{z}) \times B \hat{z}
    \end{equation}

\end{document}